### 2023-02-04 talk/talk.tex
starting a talk for Oberwolfach

%\documentclass[handout]{beamer}
\documentclass{beamer}

\usepackage{amsmath,amsfonts,amssymb,graphicx} 
\input{header.tex}
\input{header-ms.tex}
\input{theorems1.tex}

\begin{document}

\begin{frame}

\begin{center}
  {\Large\bf An iterated block particle filter for inference on coupled dynamic systems with shared and unit-specific parameters}

%% ABSTRACT: We consider inference for a collection of partially observed, stochastic, interacting, nonlinear dynamic processes. Each process is called a unit, and our primary motivation arises in biological metapopulation systems where a unit is a spatially distinct sub-population. Block particle filters are an effective tool for simulation-based likelihood evaluation for these systems, which are strongly dependent through time on a single unit and relatively weakly coupled between units. Iterated filtering algorithms can facilitate likelihood maximization for simulation-based filters. We introduce a new iterated block particle filter algorithm applicable to parameters that are either unit-specific or shared between units. We demonstrate this algorithm to carry out inference on a coupled epidemiological model for spatiotemporal measles case report data in twenty towns.

\vspace{2mm}

Edward Ionides\\
University of Michigan, Department of Statistics

\vspace{8mm}

MFO worksho on \\
Design and Analysis of Infectious Disease Studies
\\
Feb 23, 2023

\hspace{3mm}

Slides are at \url{https://ionides.github.io/talks/mfo23.pdf}
%% 8:15am EST

\vspace{8mm}

Joint work with\\
Kidus Asfaw, Ning Ning, Joonha Park, Jesse Wheeler and Aaron King

\end{center}

\end{frame}

\begin{frame}{Questions to be addressed}

  \begin{enumerate}
  \item When can we carry out full-information likelihood-based inference on a general class of spatiotemporal partially observed Markov process models?
  \item When should we want to?
  \item Is there an intersection between scientific problems of interest and the capabilities of the \texttt{spatPomp} R package?
  \end{enumerate}

  \vspace{3mm}

  We introduce the iterated block particle filter because it currently appears the most powerful algorithm available in \texttt{spatPomp}.

\end{frame}

\newcommand\challengeSep{\vspace{3mm}}

\begin{frame}{Inference challenges in population dynamics}

  \begin{enumerate}
\item Combining measurement noise and process noise.
\item Including covariates in mechanistically plausible ways.
\item  Continuous time models.
\item  Modeling and estimating interactions in coupled systems.
\item  Dealing with unobserved variables.
\item  \myemph{Modeling spatial-temporal dynamics}.
\item  \myemph{Studying population dynamics via genetic sequence data}.
  \end{enumerate}

  \vspace{3mm}

  1--6 are from Bjornstad \& Grenfell ({\it Science}, 2001).\\
  7 is from Grenfell et al ({\it Science}, 2004).\\
  1--5 are largely solved, from a methodological perspective.

\end{frame}

\begin{frame}{Example: Pre-vaccination measles in England \& Wales}

\vspace{-2mm}

\begin{center}
\includegraphics[width=8.4cm]{he10-data.pdf}

\end{center}

\vspace{-2mm}

\end{frame}

\begin{frame}{Time series data, panel data \& spatiotemporal data}

  \bi
\item Looking at one unit (town) is \myemph{time series analysis}.

  \item Joint modeling of a few units (say, 2 or 3) is \myemph{multivariate time series analysis}.

\item Analysis of many time series, without consideration of dynamic interactions, is \myemph{panel data analysis}.

\item Allowing for coupling between units, we get \myemph{spatiotemporal analysis}, which in our context is \myemph{metapopulation analysis}.

  \ei

Question: When should we avoid inference for spatiotemporal models? When do we need to consider coupling? How?

\end{frame}

\begin{frame}{Desiderata}

  \begin{itemize}
    \item We want to be able to fit arbitrary dynamic models. The limitations should be our scientific creativity and the information in the data.

    \item In practice, that means using \myemph{plug-and-play} methods which need a simulator from the model but not nice closed-form expressions for densities.

    \item We want statistically efficient inference, to extract all the information in the data.

    \item In practice, that means using likelihood-based methods.

      \item  In the time series case, iterated particle filtering (IF2) implemented in the R package \texttt{pomp} enables Masters-level statisticians to do this (\url{https://ionides.github.io/531w22/}). The science may be hard, but the statistics is becoming routine.
      \end{itemize}
  \end{frame}

\begin{frame}{Panel data}

\bi
\item To investigate epidemiological dynamics in multiple cities, one can consider each city independently, perhaps modeling a background immigration rate of infections for each city.

\item \myemph{Decoupling} leads to panel data analysis, by assumption. Iterated filtering methods extend to panel data (Breto et al, {\it Journal of the American Statistical Association}, 2019).

\item We must decide which parameters should be modeled as \myemph{shared} vs \myemph{unit-specific}.

\item The consequences of decoupling are becoming easier to study with the development of statistical inference methods for coupled systems, i.e., metapopulation dynamics.

  \ei

  \end{frame}

\begin{frame}{The curse of dimensionality}

  \bi
  \item
    Particle filter (PF) methods are effective for inference on low-dimensional nonlinear partially observed stochastic dynamic systems. They scale exponentially badly.

\item Extending the successes of particle filter methods from time series data to metapopulation data is becoming possible.

\item Algorithms under consideration:\\
  {\bf
  Bagged filters (BF, IBF)\\
  Ensemble Kalman filter (EnKF, IEnKF)\\
  Guided intermediate resampling filter (GIRF, IGIRF)\\
  Block particle filter (BPF, IBPF)\\
  }

\item Filters estimate latent states and evaluate the likelihood.
\item Each filter has an iterated version which estimates parameters by repeated filtering using stochastic parameter perturbations.

\item These algorithms are all implemented in an R package, \texttt{spatPomp}.

  \ei

\end{frame}

\begin{frame}{Filtering $U$ units of a coupled measles SEIR model}

\vspace{-3mm}

\begin{center}
\includegraphics[width=10cm]{mscale_loglik_plot-1.pdf}

\end{center}

\vspace{-2mm}

Simulated data using a gravity model with geography, demography and transmssion parameters corresponding to UK pre-vaccination measles (Ionides et al, {JASA}, 2021).

%\end{center}

\end{frame}

\begin{frame}
\frametitle{$U=40$ units for a coupled measles SEIR model}

\vspace{-2.7mm}

\begin{center}
\includegraphics[width=8cm]{slice_image_plot-1.pdf}
\end{center}

\vspace{-3mm}

{\bf A}. Simulated Susceptible-Exposed-Infected-Recovered dynamics coupled with a gravity model (log of biweekly reported cases).

{\bf B}. Measles UK pre-vaccination case reports for the 40 largest cities.

%\end{center}

\end{frame}

\begin{frame}{Parameters for the measles model}
  \bi
\item Seasonal tranmission: mean and amplitude, using school term for contact rate.
  \item Durations of latency and infectious period.
\item Initial values: fraction susceptible, latent and infectious.
\item Cohort effect: all births in an age cohort start school in September.
\item Inhomogenous mixing coefficient.
\item Measurement fraction.
\item Transport model gravity constant.
\item Dynamic noise (process overdispersion).
\item Measurement overdispersion.

\ei

  \end{frame}

\begin{frame}{More on the block particle filter}

\bi
\item BPF worked quickly, easily and reliably on our measles model filtering experiments.

\item This motivated us to develop an IBPF for parameter estimation.

\item BPF has theoretical support in some situations (Rebeschini \& Van Handel, {\it Annals of Applied Probability}, 2015).

\item BPF was independently proposed as the ``factored particle filter'' by Ng et al (2002, {\it Proc. 18th Conference on Uncertainty and Artificial Intelligence}) but not widely popularized.

\ei

\end{frame}

\begin{frame}{Particle filter (PF)}

  \begin{columns}
    \begin{column}{0.48\linewidth}
      \begin{center}
      {\bf \textcolor{blue}{Evolutionary analogy}}

      \vspace{5mm}

      {\bf Mutation}

      $\downarrow$

      {\bf Fitness}

      $\downarrow$

      {\bf Natural selection}

      \end{center}
    \end{column}
     \begin{column}{0.48\linewidth}
      \begin{center}
      {\bf \textcolor{blue}{Particle filter algorithm}}

      \vspace{5mm}

      {\bf Predict: stochastic dynamics}

      $\downarrow$

      {\bf Measurement: weight}

      $\downarrow$

      {\bf Filter: resample}
      \end{center}
    \end{column}
  \end{columns}

  \vspace{15mm}

    \begin{myitemize}
  \item PF is an evolutionary algorithm with good mathematical properties: an unbiased likelihood estimate and consistent latent state distribution.
  \end{myitemize}

\end{frame}

\begin{frame}{Block particle filter (BPF)}

  \begin{columns}
    \begin{column}{0.48\linewidth}
      \begin{center}
      {\bf \textcolor{blue}{Evolutionary analogy}}

      \vspace{5mm}

      {\bf Mutation}

      $\downarrow$

      {\bf Fitness\\
      for each chromosome}

      $\downarrow$

      {\bf Natural selection\\
      for each chromosome}

      $\downarrow$

      {\bf Recombine chromosomes}

      \end{center}
    \end{column}
     \begin{column}{0.48\linewidth}
      \begin{center}
      {\bf \textcolor{blue}{Block particle filter}}

      \vspace{5mm}

      {\bf Predict: stochastic dynamics}

      $\downarrow$

      {\bf Measurement: weight\\
      for each block}

      $\downarrow$

      {\bf Filter: resample\\
      for each block}

      $\downarrow$

      {\bf Recombine blocks}
      \end{center}
    \end{column}
  \end{columns}

  \vspace{5mm}

    \begin{myitemize}
    \item Blocks in BPF allow recombination (reassortment of chromosomes in sexual reproduction) in the evolutionary analogy.
    \item Blocks are a partition of the metapopulation units. Our experiments suggest treating each sub-population (i.e., city) as a block.

  \end{myitemize}

\end{frame}

\begin{frame}

\frametitle{Measles likelihood slices for coupling parameter, $G$}

\vspace{-5mm}

\begin{columns}[T] % align columns
\begin{column}{.55\textwidth}
  \includegraphics[width=6cm]{slice_combined_plot-1.pdf}
\end{column}
\begin{column}{.52\textwidth}

  \vspace{6mm}

  Simulating $15$ year of data from $U=40$ cities for the measles model.
  Slice likelihood, varying $G$ with other paramters fixed at the truth.

  \vspace{3mm}

  {\bf A}. Evaluation using adapted bagged filter (ABF).

    \vspace{3mm}

  {\bf B}. Evaluation using block particle filter (BPF).

    \vspace{3mm}

  {\bf C}. Evaluation using EnKF.

  \vspace{3mm}

Results from Ionides et al (2021, {\it JASA}). We computed a slice due to lack of good optimization algorithms to compute a profile.

\end{column}
\end{columns}

\end{frame}

\begin{frame}{Scalability needed for practical inference}

Large numbers of parameters
  \begin{itemize}
  \item  Initial conditions will typically have to be estimated for each unit.
  \item Various dynamic parameters and measurement parameters (e.g., reporting rate) may also need to be unit-specific to obtain a statistical fit to the data.
\item Working with hundreds of estimated parameters raises additional challenges on top of the high-dimensional coupled dynamics.
  \end{itemize}

\vspace{5mm}

A moderate numbers of spatial units is enough to open new possibilities.

  \begin{itemize}
  \item As soon as dimension exceeds capabilities of a particle filter (say, $U=5$) we are in situations where likelihood-based inference for general models has been inaccessible.

  \item 10-100 coupled units is our target problem size.

  \item Larger problems will need numerical approximations (e.g., EnKF). Exact Monte Carlo methods help study the effect of these approximations.

  \end{itemize}

  \end{frame}

\begin{frame}{An iterated block particle filter for parameter estimation}

  \begin{center}

  \includegraphics[trim={0 0 0 10mm},clip,width=9cm]{IBPF_workflow.pdf}

  \end{center}

\end{frame}

\begin{frame}

\vspace{-6mm}

\begin{columns}[T] % align columns
\begin{column}{.75\textwidth}
  \includegraphics[width=6.8cm]{ibpf/search_diagnostics-1.pdf}
\end{column}

\hspace{-15mm}
\begin{column}{.35\textwidth}
  \vspace{10mm}

$\bullet$    IBPF applied to simulations and data.

    \vspace{5mm}

$\bullet$    Multiple searches from random starting points.

    \vspace{5mm}

$\bullet$    Top 25\% of searches are subsequently refined.

  \end{column}
\end{columns}

\end{frame}

\begin{frame}

\begin{columns}[T] % align columns
\begin{column}{.75\textwidth}
  \includegraphics[width=7.5cm]{ibpf/sim_search_boxplot-1.pdf}

    \includegraphics[width=7.5cm]{ibpf/data_search_boxplot-1.pdf}

  \end{column}

\hspace{-15mm}
\begin{column}{.35\textwidth}
  \vspace{5mm}

$\bullet$    IBPF applied to simulations (top panel) and data (bottom panel).

    \vspace{5mm}

$\bullet$    Multiple search refinements.

    \vspace{5mm}

    (A) Mostly shared parameters.

    \vspace{5mm}

    (B) All unit-specific parameters

    \vspace{5mm}

    (C) all unit-specific, without coupling

  \end{column}
\end{columns}

\end{frame}

\begin{frame}
\frametitle{Filtering $U$-dimensional correlated Brownian motion}

\vspace{-3mm}

\begin{center}
\includegraphics[width=10cm]{bm_alt_plot-1.pdf}

\vspace{-1mm}

$\cov\big(X_{\unit,\time}-X_{\unit,\time-1},X_{\altUnit,\time}-X_{\altUnit,\time-1}\big) \sim 0.4^{|\unit-\altUnit|}_{}$

\end{center}

\end{frame}

\begin{frame}
\frametitle{Filtering $U$ units of Lorenz 96 toy atmospheric model}

\vspace{-3mm}

\begin{center}
\includegraphics[width=10cm]{lz_loglik_plot-1.pdf}

\vspace{-1mm}

$dX_{\unit}(t) = \big \{  X_{\unit-1}(t) \big(X_{\unit+1}(t) - X_{\unit-2}(t)\big) - X_{\unit}(t) + F \big\} dt + \sigma \, dB_{\unit}(t)$

\end{center}

\end{frame}

\begin{frame}{Future work}

  \newcommand\futuresep{\vspace{3mm}}

  \begin{myitemize}
  \item We are reaching to the point where we can carry out likelihood-based inference for a flexible class of metapopulation models for measles.
Flexibility supports generation and testing of scientific hypotheses.

    \futuresep

%  \item Measles was previously a motivating model system for POMP methods for single populations.

    \futuresep

  \item Many systems in ecology, epidemiology and elsewhere could be studied in a SpatPOMP framework. Including microbiomes?

     \futuresep

   \item Modeling and inference for nonlinear stochastic dynamics is hard. But, if you can't build a quantitative statistical model then you don't understand it and you can't control it?

\end{myitemize}

\end{frame}

\nocite{bjornstad01,grenfell04,breto19,rebeschini15,ng02,ionides21}

\begin{frame}[allowframebreaks]
\frametitle{References}
\bibliographystyle{apalike}
\bibliography{bib}
\end{frame}

%%% extra material

\end{document}