### 2023-02-23 talk/talk.tex
talk for mfo23

%\documentclass[handout]{beamer}
\documentclass{beamer}

\usepackage{amsmath,amsfonts,amssymb,graphicx} 
\input{header.tex}
\input{header-ms.tex}
\input{theorems1.tex}

\begin{document}

\begin{frame}

\begin{center}
  {\Large\bf An iterated block particle filter for inference on coupled dynamic systems with shared and unit-specific parameters}

%% ABSTRACT: We consider inference for a collection of partially observed, stochastic, interacting, nonlinear dynamic processes. Each process is called a unit, and our primary motivation arises in biological metapopulation systems where a unit is a spatially distinct sub-population. Block particle filters are an effective tool for simulation-based likelihood evaluation for these systems, which are strongly dependent through time on a single unit and relatively weakly coupled between units. Iterated filtering algorithms can facilitate likelihood maximization for simulation-based filters. We introduce a new iterated block particle filter algorithm applicable to parameters that are either unit-specific or shared between units. We demonstrate this algorithm to carry out inference on a coupled epidemiological model for spatiotemporal measles case report data in twenty towns.

\vspace{2mm}

Edward Ionides\\
University of Michigan, Department of Statistics

\vspace{8mm}

MFO workshop on \\
Design and Analysis of Infectious Disease Studies
\\
Feb 23, 2023

\hspace{3mm}

Slides are at \url{https://ionides.github.io/talks/mfo23.pdf}
%% 8:15am EST

\vspace{8mm}

Joint work with
Patricia Ning, Jesse Wheeler, Kidus Asfaw, Jifan Li, Joonha Park, Aaron King

\end{center}

\end{frame}

\begin{frame}{Questions to be addressed}

  \begin{enumerate}
  \item When can we carry out full-information likelihood-based inference on a general class of spatiotemporal partially observed Markov process models? (Of particular interest are metapopulation models.)

    \vspace{2mm}

  \item Is there an intersection between scientific problems of interest and the capabilities of the \texttt{spatPomp} R package?

        \vspace{2mm}

\item We introduce the {\bf iterated block particle filter} because it is currently the most powerful algorithm available in \texttt{spatPomp}.

  \vspace{2mm}

\item Bonus question: How do we know if our model is statistically adequate, or needs more work?

    \end{enumerate}
\end{frame}

\newcommand\challengeSep{\vspace{3mm}}

\begin{frame}{Inference challenges in population dynamics}

  \begin{enumerate}
\item Combining measurement noise and process noise.
\item Including covariates in mechanistically plausible ways.
\item  Continuous time models.
\item  Modeling and estimating interactions in coupled systems.
\item  Dealing with unobserved variables.
\item  \myemph{Modeling spatiotemporal dynamics}.
\item  Studying population dynamics via genetic sequence data.
  \end{enumerate}

  \vspace{4mm}

  1--5 are largely solved, from a methodological perspective.\\
  6 is our immediate topic.\\
  7 is exciting but not the focus of this talk.

  \vspace{4mm}

  \resizebox{12cm}{!}{Reviews: Bjornstad \& Grenfell ({\it Science}, 2001); Grenfell et al ({\it Science}, 2004)}

\end{frame}

\begin{frame}{Pre-vaccination weakly measles in England \& Wales}

  %% full weekly data now available:
  %% Korevaar, Hannah, C. Jessica Metcalf, and Bryan T. Grenfell. "Structure, space and size: competing drivers of variation in urban and rural measles transmission." Journal of the Royal Society Interface 17, no. 168 (2020): 20200010.
  %% https://royalsocietypublishing.org/doi/pdf/10.1098/rsif.2020.0010
  %% data at https://rs.figshare.com/collections/Supplementary_material_from_Structure_space_and_size_competing_drivers_of_variation_in_urban_and_rural_measles_transmission_/5036567/1

\vspace{-2mm}

\begin{center}
\includegraphics[width=8.4cm]{he10-data.pdf}

\end{center}

\vspace{-2mm}

\end{frame}

\begin{frame}{Time series data, panel data \& spatiotemporal data}

  \bi
\item Looking at one unit (town) is \myemph{time series analysis}.

  \item Joint modeling of a few units (say, 2 or 3) is \myemph{multivariate time series analysis}.

\item Analysis of many time series, without consideration of dynamic interactions, is \myemph{panel data analysis}.

\item Allowing for coupling between units, we get \myemph{spatiotemporal analysis}, which in our context is \myemph{metapopulation analysis}.

  \ei

%\vspace{3mm}

%Question: When should we avoid inference for spatiotemporal models? When do we need to consider coupling? How?

\end{frame}

\begin{frame}{Desiderata}

  \begin{itemize}
    \item Consideration of arbitrary dynamic models. The limitations should be our scientific creativity and the information in the data.

      \vspace{2mm}

     Hence, \myemph{plug-and-play} methods which need a simulator from the model but not nice closed-form expressions for densities.

     \vspace{2mm}

    \item Statistically efficient inference, to extract all the information in the data.

      \vspace{2mm}

    Hence, \myemph{likelihood-based} methods.

      \end{itemize}
  \end{frame}

\begin{frame}{Fitting mechanistic models to time series}

    \vspace{8mm}

    \begin{itemize}
    \item  Iiterated particle filtering via \texttt{mif2} in the R package \texttt{pomp} enables Masters-level statisticians to do plug-and-play likelihood-based inference:

      \vspace{2mm}

      \url{https://ionides.github.io/531w22/}

      \vspace{5mm}
      \item
      The science may be hard, but the statistics is becoming routine.

    % examples: COVID etc, bison in Yellowstone National Park, volatility in the stock market, gun purchases, human proximal small intestine pH

    \end{itemize}
\end{frame}

\begin{frame}{Fitting mechanistic models to panel data}

\bi
\item To investigate epidemiological dynamics in multiple cities, one can consider each city independently, perhaps modeling a background immigration rate of infections for each city.

  \vspace{2mm}

\item \myemph{Decoupling} leads to panel data analysis, by assumption.

  \vspace{2mm}

  \item Iterated filtering methods extend to panel data (\texttt{panelPomp}, Breto et al, {\it JASA}, 2019).

    \vspace{2mm}

\item We must decide which parameters should be modeled as \myemph{shared} (i.e., the same for each unit) vs \myemph{unit-specific}.

%\item The consequences of decoupling are becoming easier to study with the development of statistical inference methods for coupled systems, i.e., metapopulation dynamics.

  \ei

  \end{frame}

\begin{frame}{The curse of dimensionality}

  \bi
  \item
    Particle filter (PF) methods fail for high-dimensional systems. They scale exponentially badly.

    \vspace{2mm}

  \item Algorithms with improved scalability include:\\

    \vspace{1mm}

  {\bf
  Bagged filters (BF, IBF)\\
  Ensemble Kalman filter (EnKF, IEnKF)\\
  Guided intermediate resampling filter (GIRF, IGIRF)\\
  Block particle filter (BPF, IBPF)\\
  }

      \vspace{2mm}

\item Filters estimate latent states and evaluate the likelihood.

    \vspace{2mm}

  \item Iterated filters estimate parameters using stochastic parameter perturbations.

        \vspace{2mm}

\item These algorithms are all implemented in the \texttt{spatPomp} R package.

  \ei

\end{frame}

\begin{frame}{Filtering $U$ units of a coupled measles SEIR model}

\vspace{-1mm}

\begin{center}
\includegraphics[width=10cm]{mscale_loglik_plot-1.pdf}

\end{center}

\vspace{-4mm}

Simulated data using a gravity model with geography, demography and transmssion parameters corresponding to UK pre-vaccination measles (Ionides et al, {JASA}, 2021).

%\end{center}

\end{frame}

\begin{frame}
\frametitle{$U=40$ units for a coupled measles SEIR model}

\vspace{-2.7mm}

\begin{center}
\includegraphics[width=8cm]{slice_image_plot-1.pdf}
\end{center}

\vspace{-3mm}

{\bf A}. Simulated Susceptible-Exposed-Infected-Recovered dynamics coupled with a gravity model (log of biweekly reported cases).

{\bf B}. Measles UK pre-vaccination case reports for the 40 largest cities.

%\end{center}

\end{frame}

\begin{frame}{Parameters for the measles model}
  \bi
\item Seasonal transmission: mean and amplitude, using school term for contact rate.
  \item Durations of latency and infectious period.
\item Initial values: fraction susceptible, latent and infectious.
\item Cohort effect: all births in an age cohort start school in September.
\item Inhomogenous mixing exponent, $I^{\alpha}$.
\item Measurement fraction.
\item Population movement model: gravity constant.
\item Dynamic noise: process overdispersion.
\item Measurement overdispersion.

\ei
\vspace{3mm}
(Adding gravity coupling to the model of He et al, {\it JRSI}, 2010)
%\vspace{5mm}
%Plug-and-play methods let you add or subtract features

  \end{frame}

\begin{frame}{More on the block particle filter (BPF)}

\bi
\item BPF worked quickly, easily and reliably on our measles model filtering experiments.

        \vspace{2mm}

\item BPF has theoretical support in some situations (Rebeschini \& Van Handel, {\it Annals of Applied Probability}, 2015).

        \vspace{2mm}

      \item This motivated us to develop an IBPF for parameter estimation.

        \vspace{2mm}

      \item IBPF has theoretical guarantees similar to BPF (Ning \& Ionides, {\it arXiv}, 2023).

        \vspace{2mm}

\item BPF was independently proposed as the ``factored particle filter'' by Ng et al (2002).

\ei

\end{frame}

\begin{frame}{Particle filter (PF)}

  \begin{columns}
    \begin{column}{0.48\linewidth}
      \begin{center}
      {\bf \textcolor{blue}{Evolutionary analogy}}

      \vspace{5mm}

      {\bf Mutation}

      $\downarrow$

      {\bf Fitness}

      $\downarrow$

      {\bf Natural selection}

      \end{center}
    \end{column}
     \begin{column}{0.48\linewidth}
      \begin{center}
      {\bf \textcolor{blue}{Particle filter algorithm}}

      \vspace{5mm}

      {\bf Predict: stochastic dynamics}

      $\downarrow$

      {\bf Measurement: weight}

      $\downarrow$

      {\bf Filter: resample}
      \end{center}
    \end{column}
  \end{columns}

  \vspace{15mm}

    \begin{myitemize}
  \item PF is an evolutionary algorithm with good mathematical properties: an unbiased likelihood estimate and consistent latent state distribution.
  \end{myitemize}

\end{frame}

\begin{frame}{Block particle filter (BPF)}

  \begin{myitemize}
    \item Blocks are a partition of the metapopulation units.
    \item For measles, we use each city as a block.
  \end{myitemize}

  \begin{columns}
    \begin{column}{0.48\linewidth}
      \begin{center}
      {\bf \textcolor{blue}{Evolutionary analogy}}

      \vspace{5mm}

      {\bf Mutation}

      $\downarrow$

      {\bf Fitness\\
      for each chromosome}

      $\downarrow$

      {\bf Natural selection\\
      for each chromosome}

      $\downarrow$

      {\bf Recombine chromosomes}

      \end{center}
    \end{column}
     \begin{column}{0.48\linewidth}
      \begin{center}
      {\bf \textcolor{blue}{Block particle filter}}

      \vspace{5mm}

      {\bf Predict: stochastic dynamics}

      $\downarrow$

      {\bf Measurement: weight\\
      for each block}

      $\downarrow$

      {\bf Filter: resample\\
      for each block}

      $\downarrow$

      {\bf Recombine blocks}
      \end{center}
    \end{column}
  \end{columns}

  \vspace{5mm}

    \begin{myitemize}
    \item Blocks are segments of the full state which can be reassorted between particles at the resampling step.

\end{myitemize}

\end{frame}

\begin{frame}

\frametitle{Measles likelihood slices for coupling parameter, $G$}

\vspace{-5mm}

\begin{columns}[T] % align columns
\begin{column}{.55\textwidth}
  \includegraphics[width=6cm]{slice_combined_plot-1.pdf}
\end{column}

\hspace{-5mm}
\begin{column}{.52\textwidth}

  \vspace{6mm}

  Simulating $15$ year of data from $U=40$ cities.
  Slice likelihood, varying $G$ with other parameters fixed.\\
   (Ionides et al, {\it JASA}, 2021).

  \vspace{3mm}

  {\bf A}. Evaluation using adapted bagged filter (ABF).

    \vspace{3mm}

  {\bf B}. Evaluation using block particle filter (BPF).

    \vspace{3mm}

  {\bf C}. Evaluation using EnKF.

  \vspace{3mm}
  A much easier task than computing a profile likelihood for real data.

\end{column}
\end{columns}

\end{frame}

\begin{frame}{Comments on the Ensemble Kalman Filter (EnKF)}

  \begin{itemize}
  \item EnKF is more dependent on approximate Gaussianity than is sometimes supposed.

    \vspace{2mm}

  \item The Gaussian-inspired update rule is similar to the extended Kalman filter (EKF), which has largely been superseded by particle filter methods for low-dimensional nonlinear biological dynamics.

    \vspace{2mm}

\item Simple systems can defeat EnKF: the linear Gaussian update is helpless when data inform the conditional variance rather than the conditional mean.

  \vspace{2mm}

    \item Big systems need computationally tractable analysis. EnKF may sometimes be the best solution available, but be aware of its limitations.

  \end{itemize}

\end{frame}

\begin{frame}{Scalability needed for practical inference}

Large numbers of parameters
  \begin{itemize}
  \item  Initial conditions will typically have to be estimated for each unit.
  \item Various dynamic parameters and measurement parameters (e.g., reporting rate) may also need to be unit-specific to obtain a statistical fit to the data.
\item Working with hundreds of estimated parameters raises additional challenges on top of the high-dimensional coupled dynamics.
  \end{itemize}

\vspace{5mm}

A moderate numbers of spatial units is enough to open new possibilities.

  \begin{itemize}
  \item As soon as dimension exceeds capabilities of a particle filter (say, $U=5$) we are in situations where likelihood-based inference for general models has been inaccessible.

  \item 5-500 coupled units is our target problem size.

  \item Larger problems will need numerical approximations (e.g., EnKF).

  \end{itemize}

  \end{frame}

\begin{frame}{An iterated block particle filter for parameter estimation}

  \begin{center}

  \includegraphics[trim={0 0 0 10mm},clip,width=9cm]{IBPF_workflow.pdf}

  \end{center}

\end{frame}

\begin{frame}

%%% A: 9 shared parameters; 4 unit-specific
%%% B: all unit-specific

\vspace{-6mm}

\begin{columns}[T] % align columns
\begin{column}{.75\textwidth}
  \includegraphics[width=6.8cm]{ibpf/search_diagnostics-1.pdf}
\end{column}

\hspace{-15mm}
\begin{column}{.35\textwidth}
  \vspace{10mm}

$\bullet$    IBPF applied to simulations and data.

    \vspace{5mm}

$\bullet$    Multiple searches from random starting points.

    \vspace{5mm}

$\bullet$    Top 25\% of searches are subsequently refined.

  \end{column}
\end{columns}

\end{frame}

\begin{frame}

\begin{columns}[T] % align columns
\hspace{-4mm}\begin{column}{.75\textwidth}
  \includegraphics[width=7.5cm]{ibpf/sim_search_boxplot-1.pdf}

    \includegraphics[width=7.5cm]{ibpf/data_search_boxplot-1.pdf}

  \end{column}

\hspace{-18mm}
\begin{column}{.4\textwidth}
  \vspace{5mm}

  {\bf IBPF for measles, 20 cities}\\

  \vspace{1mm}

  $\bullet$ simulations (top panel)\\
  dash line: likelihood at truth

  \vspace{2mm}

  $\bullet$ data (bottom panel)\\
  dash line: MLE for case C.

    \vspace{5mm}

%$\bullet$    Multiple search refinements.

%    \vspace{5mm}

    (A) 9 shared parameters, $4\times 20$ unit-specific parameters

    \vspace{5mm}

    (B) all parameters unit-specific

    \vspace{5mm}

    (C) all unit-specific, with constant immigration instead of coupling

  \end{column}
\end{columns}

\end{frame}

\begin{frame}{Inferences from the previous slide}

%%% Wilks' Theorem: improvement on truth is about (1/2) chisq(p)

\begin{itemize}
  \item
IBPF can work for simulated data, since it finds likelihoods higher than the truth when the model is well-specified

  \vspace{5mm}

  \item
    When coupled by a gravity model, simulated data show evidence against a constant immigration model.

    \vspace{5mm}

\item 
Data do not favor this gravity model over a constant immigration model.

%%% An underwhelming sounding result, but a necessary step to make progress,
%%% and previously we did not know how to do this.

\end{itemize}

\end{frame}

\begin{frame}{He et al (2010) estimates. $r$ is rank correlation with $N_{1950}$}

\resizebox{12cm}{!}{
\begin{tabular}{l|r|r|r|r|r|r|r|r|r|r}
\hline
  & $N_{1950}$ & $R_0$ & IP & LP & $\alpha$ & $a$ & $\iota$ & $\psi$ & $\rho$ & $\sigma_{SE}$\\
\hline
Halesworth & 2200 & 33.00 & 2.30 & 7.90 & 0.95 & 0.38 & 0.0091 & 0.64 & 0.75 & 0.075\\
\hline
Lees & 4200 & 30.00 & 2.10 & 8.50 & 0.97 & 0.15 & 0.0310 & 0.68 & 0.61 & 0.080\\
\hline
Mold & 6400 & 21.00 & 1.80 & 5.90 & 1.00 & 0.27 & 0.0140 & 2.90 & 0.13 & 0.054\\
\hline
Dalton in Furness & 11000 & 28.00 & 2.00 & 5.50 & 0.99 & 0.20 & 0.0390 & 0.82 & 0.46 & 0.078\\
\hline
Oswestry & 11000 & 53.00 & 2.70 & 10.00 & 1.00 & 0.34 & 0.0300 & 0.48 & 0.63 & 0.070\\
\hline
Northwich & 18000 & 30.00 & 3.00 & 8.50 & 0.95 & 0.42 & 0.0600 & 0.40 & 0.80 & 0.086\\
\hline
Bedwellty & 29000 & 25.00 & 3.00 & 6.80 & 0.94 & 0.16 & 0.0400 & 0.95 & 0.31 & 0.061\\
\hline
Consett & 39000 & 36.00 & 2.70 & 9.10 & 1.00 & 0.20 & 0.0730 & 0.41 & 0.65 & 0.071\\
\hline
Hastings & 66000 & 34.00 & 5.40 & 7.00 & 1.00 & 0.30 & 0.1900 & 0.40 & 0.70 & 0.096\\
\hline
Cardiff & 240000 & 34.00 & 3.10 & 9.90 & 1.00 & 0.22 & 0.1400 & 0.27 & 0.60 & 0.054\\
\hline
Bradford & 290000 & 32.00 & 3.40 & 8.50 & 0.99 & 0.24 & 0.2400 & 0.19 & 0.60 & 0.045\\
\hline
Hull & 300000 & 39.00 & 5.50 & 9.20 & 0.97 & 0.22 & 0.1400 & 0.26 & 0.58 & 0.064\\
\hline
Nottingham & 310000 & 23.00 & 3.70 & 5.70 & 0.98 & 0.16 & 0.1700 & 0.26 & 0.61 & 0.038\\
\hline
Bristol & 440000 & 27.00 & 4.90 & 6.20 & 1.00 & 0.20 & 0.4400 & 0.20 & 0.63 & 0.039\\
\hline
Leeds & 510000 & 48.00 & 11.00 & 9.50 & 1.00 & 0.27 & 1.2000 & 0.17 & 0.67 & 0.078\\
\hline
Sheffield & 520000 & 33.00 & 6.40 & 7.20 & 1.00 & 0.31 & 0.8500 & 0.18 & 0.65 & 0.043\\
\hline
Manchester & 700000 & 33.00 & 6.90 & 11.00 & 0.96 & 0.29 & 0.5900 & 0.16 & 0.55 & 0.055\\
\hline
Liverpool & 800000 & 48.00 & 9.80 & 7.90 & 0.98 & 0.30 & 0.2600 & 0.14 & 0.49 & 0.053\\
\hline
Birmingham & 1100000 & 43.00 & 12.00 & 8.50 & 1.00 & 0.43 & 0.3400 & 0.18 & 0.54 & 0.061\\
\hline
London & 3400000 & 57.00 & 13.00 & 13.00 & 0.98 & 0.55 & 2.9000 & 0.12 & 0.49 & 0.088\\
\hline
$\qquad r$ & 1 & 0.46 & \cellcolor{yellow!50} 0.95 & 0.32 & 0.11 & 0.30 & 0.9300 & -0.93 & -0.20 & -0.330\\
\hline
\end{tabular}
}

\end{frame}

\begin{frame}{Cholera in Haiti, 2010-2019}

% https://web.archive.org/web/20151106110552/http://www.ihsi.ht/pdf/projection/Estimat_PopTotal_18ans_Menag2015.pdf
%Map	Department	Capital	Area (km2)	Population	Density (Pop./km2)	Planning Region
%1	Artibonite	Gonaïves	4,987	1,727,524	350	Central
%2	Centre	        Hinche	        3,487	746,236	        210	Central
%3	Grand'Anse	Jérémie	        1,912	468,301	        240	South
%4	Nippes	        Miragoâne	1,268	342,525	        270	South
%5	Nord	        Cap-Haïtien	2,115	1,067,177	500	North
%6	Nord-Est	Fort-Liberté	1,623	393,967	        240	North
%7	Nord-Ouest	Port-de-Paix	2,103	728,807	        350	Central
%8	Ouest	        Port-au-Prince	4,983	4,029,705	810	West
%9	Sud-Est	        Jacmel	        2,034	632,601	        310	West
%10	Sud	        Les Cayes	2,654	774,976	        290	South

  % Hurricane Matthew: October 2016

\begin{columns}[T] % align columns

%\hspace{-15mm}
\begin{column}{.35\textwidth}
  \vspace{2mm}

  $\bullet$  820,000 reported cases

  \vspace{4mm}

  $\bullet$ nearly 10,000 deaths

    \vspace{4mm}

$\bullet$  Population of 10,000,000 in 10 d\'{e}partements

    \vspace{4mm}

    $\bullet$  Lee et al (2020) developed 4 models to guide vaccination plans:

    %%% Elizabeth Lee, Johns Hopkins

    \vspace{1mm}

    {\bf 1}. stochastic, national \\
    {\bf 2}. ODE, spatial \\
    {\bf 3}. stochastic, spatial\\
    {\bf 4}. agent based

\end{column}

\begin{column}{.62\textwidth}
  \includegraphics[width=7.5cm]{haiti/haiti.png}

  \vspace{-3mm}

  \begin{center}
    {  \tiny \url{https://doi.org/10.5194/nhess-20-471-2020}}
  \end{center}

\end{column}

\end{columns}

  \vspace{4mm}

   $\bullet$ Wheeler et al (2023) continued the data analysis

\end{frame}

\begin{frame}{Haiti cholera weekly reports 2010-2019 (log scale)}
\includegraphics[width=12cm]{haiti/Plot_Reported_Cases-1.pdf}
\end{frame}

\begin{frame}{Log-likelihoods of models and an ARMA benchmark}

  \begin{center}
\begin{tabular}{|l|rc|rc|rc|}
\hline
& Model~1& (p) & Model~2 & (p) & Model~3 & (p)
\\
\hline
Wheeler et al
& \cellcolor{yellow!50} -2731.3 & \cellcolor{yellow!50}(15)
& -21957.3 & (6)
& \cellcolor{yellow!50} -17850.4 & \cellcolor{yellow!50}(35)
\\
Lee et al
& -3030.9 & (20)
& -29367.4 & (6)
& -31840.8 & (29)
\\
Log-ARMA(2,1) \hspace{-2.5mm}
& -2803.7 & (4)  %% -0.5*5615.3+4
& \cellcolor{yellow!50} { -18027.0} & \cellcolor{yellow!50} (40)  %% -0.5*36133.9+40
& -18027.0 & (40)
\\
\hline
\end{tabular}
  \end{center}

  \vspace{3mm}

  Model 1 (stochastic, aggregated): \\
  We add process overdispersion, and beat the ARMA benchmark.

  \vspace{5mm}

  Model 2 (deterministic, spatial): \\
  We use a log-normal measurement model. ODE models fall far short of the ARMA benchmark.

    \vspace{5mm}

    Model 3 (stochastic, spatial): \\
    We fit the model using IBPF, and beat the ARMA benchmark.

\end{frame}

\begin{frame}{Evidence for a decreasing transmission rate in Model 1}
\includegraphics[width=12cm]{haiti/Beta_trend_Figure-1.pdf}
\end{frame}

% From Li et al (2020) page S-14:
% Ruiyun Li, Imperial (Oslo as of 2021)
% According to official report, around 5 million people left Wuhan city before January 23. The total population in Wuhan city after January 23 (when travel restrictions were imposed) is therefore around 6 million.

\begin{frame}{COVID in 375 cities in China, 10-23 January, 2020}

  \begin{itemize}
  \item Metapopulation data were used to infer the fraction of asymptomatic cases and their contagiousness
    (Li et al, {\it Science}, May 2020).
  \item SEIR model with asymptomatics, reporting delay, and coupling based on cell phone data.
  \item Inference using iterated EnKF.

    \vspace{5mm}
   \item
    The time interval predates Wuhan travel restrictions.
%% see README for R code
   \item Total cases 801, Wuhan 454, Chongqing 27, Beijing 26.
   \item{} 259 cities with 0 cases, 99 with 1--5 cases, 17 with $>5$ cases.
   \item{} 0 cases reported for all cities during Jan 10--15.

\end{itemize}

\end{frame}

\begin{frame}{Top: data. Middle: IBPF fit. Bottom: IEnKF fit} 
\includegraphics[width=12cm]{covid/3-panel.pdf}
\end{frame}

\begin{frame}{Technical issues with EnKF}

  \begin{itemize}
  \item EnKF is based on a continuous Gaussian approximation.
  \item Log-likelihoods with respect to counting measure are never positive, and cannot properly be compared with log-likelihoods corresponding to continuous densities.
    \item For data with many zeros, the unbounded EnKF likelihood can substantially bias the MLE.
    \item Adapting EnKF for count data is non-trivial (Katzfuss et al, {\it JASA}, 2019).
    \item Li et al avoided the issue by using a ``heuristic observation error variance'' within EnKF. The corresponding ``log-likelihood'' is not the log-likelihood of any model. In particular, it cannot validly be compared againt simple statistical benchmarks.
  \item Data and models will never be perfect. We want to reveal, rather than conceal, limitations.
  \end{itemize}

  \end{frame}

\begin{frame}{Future work}

  \newcommand\futuresep{\vspace{5mm}}

  \begin{myitemize}

  \item We anticipate continued progress in modeling and inference for nonlinear spatiotemporal partially observed stochastic dynamic systems.

    \futuresep

  \item The \texttt{spatPomp} package facilitates data analysis.

    \futuresep

    \item Questions and comments?

\end{myitemize}

\end{frame}

\begin{frame}
\frametitle{Filtering $U$-dimensional correlated Brownian motion}

\vspace{-3mm}

\begin{center}
\includegraphics[width=10cm]{bm_alt_plot-1.pdf}

\vspace{-1mm}

$\cov\big(X_{\unit,\time}-X_{\unit,\time-1},X_{\altUnit,\time}-X_{\altUnit,\time-1}\big) \sim 0.4^{|\unit-\altUnit|}_{}$

\end{center}

\end{frame}

\begin{frame}
\frametitle{Filtering $U$ units of Lorenz 96 toy atmospheric model}

\vspace{-3mm}

\begin{center}
\includegraphics[width=10cm]{lz_loglik_plot-1.pdf}

\vspace{-1mm}

$dX_{\unit}(t) = \big \{  X_{\unit-1}(t) \big(X_{\unit+1}(t) - X_{\unit-2}(t)\big) - X_{\unit}(t) + F \big\} dt + \sigma \, dB_{\unit}(t)$

\end{center}

\end{frame}

\nocite{asfaw23arxiv,bjornstad01,breto19,grenfell04,he10,katzfuss19,ionides21,ionides22,lee20,li20,ng02,ning23-ibpf,park20,rebeschini15,wheeler23}

\begin{frame}[allowframebreaks]
\frametitle{References}
\bibliographystyle{apalike}
\bibliography{bib}
\end{frame}

%%% extra material

\end{document}